---
title: "Writing Data Frame & Plot Functions in R"
format: html
number-sections: true
number-depth: 2
bibliography: references.bib
editor: source
execute: 
  echo: true
  include: true
---

```{r}
#| include: false
#| echo: false

library(tidyverse)
library(lubridate)

```

The second half of this week's coursework focuses on writing functions that work with *data frames*.

By the end of the week you should have a grasp of:

-   What the "embracing" (`{{{ }}}`) operator is

-   How to use the `{{{ }}}` operator in data frame functions

-   Why we need to use the `{{{ }}}` operator when writing data frame functions

------------------------------------------------------------------------

#### `r emo::ji("play button")` Watch Videos: 20 minutes

#### `r emo::ji("book")` Readings: 60-75 minutes

#### `r emo::ji("check")` Preview Activities: 1 (broken into two parts)

------------------------------------------------------------------------

# Writing Data Frame Functions in R

### `r emo::ji("book")` [Required Reading: *R4DS* -- Functions](https://r4ds.hadley.nz/functions#data-frame-functions)

::: callout-caution
# Read only Section 3 (Data frame functions)!
:::

## Tidy Evaluation

Writing functions that work with data frames **and** call on the functions we've become used to (e.g., `filter()`, `select()`, `summarise()`) requires we learn about tidy evaluation. To write these functions you will need to know, at a high level, whether the function you are trying to incorporate uses data masking or tidy selection.

At a high level, data masking is used in functions like `arrange()`, `filter()`, and `summarize()` that compute with variables. Whereas, tidy selection is used for functions like `select()` and `rename()` that select variables.

Your intuition about which functions use tidy evaluation should be good for many of these functions. If you can input `c(var1, var2, var3)` into the function (e.g., `select(mtcars, c(vs, am, gear))`), then the function uses tidy selection! If you **cannot** input `c(var1, var2, var3)` into the function, then the function is performing computations on the data and uses data masking.

If you are interested in learning more about tidy evaluation, I would highly recommend:

-   this [video by Jenny Bryan](https://youtu.be/2BXPLnLMTYo?si=ry4tHOD77DWagvxn)
    -   I do want to note that this video is from 2019 and some things have changed since then. Namely, we used to need to use the `enquo()` function to inject\
        variable names into `dplyr` functions, whereas we now use embracing `{{}}`. 🤗
-   this [vignette for tidy evaluation with `dplyr`](https://dplyr.tidyverse.org/articles/programming.html#:~:text=arrange()%20%2C%20count()%20%2C%20filter,my_variable%20not%20df%24my_variable%20)

### `r emo::ji("check")` **Check-in 7.2: Writing Data Frame Functions**

::: callout-note
# Data Structure

Note for Questions 1 & 2, I am assuming the `flight` data has a similar structure to the `nycflights` data from the openintro R package.

```{r}
head(openintro::nycflights)
```
:::

**Question 1**: Fill in the code below to write a function that finds all flights that were cancelled or delayed by more than a user supplied number of hours:

```{r}
#| eval: false

filter_severe <- function(df, hours) { 
  df |> 
    filter(dep_delay _____)
}


nycflights |> 
  filter_severe(hours = 2)
```

**Question 2**: Fill in the code below to write a function that converts the user supplied variable that uses clock time (e.g., `dep_time`, `arr_time`, etc.) into a decimal time (i.e. hours + (minutes / 60)).

```{r}
#| eval: false

standardize_time <- function(df, time_var) {
  df |> 
    # Times are stored as 2008 for 8
    mutate( {{ time_var }} := 
              as.numeric(
                ## Grab first two numbers for hour
                str_sub(
                  {{ time_var }}, 
                  start = 1, 
                  end = 2)
                ) +  
              as.numeric(
                ## Grab second two numbers for minutes
                str_sub(
                  {{ time_var }}, 
                  start = 3, 
                  end = 4)
                ) / 60
            )
  
}

nycflights |> 
  standardize_time(arr_time)
```

**Question 3**: For each of the following functions determine if the function uses data-masking or tidy-selection:

-   `distinct()`
-   `count()`
-   `group_by()`
-   `select()`
-   `rename_with()`
-   `across()`

# Writing Plotting Functions in R

### `r emo::ji("book")` [Required Reading: *R4DS* -- Functions](https://r4ds.hadley.nz/functions#plot-functions)

::: callout-caution
# Read only Section 4 (Data frame functions)!
:::

### `r emo::ji("check")` **Check-in 7.2: Writing Data Frame Functions**

**Question 4**: Fill in the code below to build a rich plotting function which:

-   draws a scatterplot given dataset and x and y variables,
-   adds a line of best fit (i.e. a linear model with no standard errors)
-   add a title.

```{r}
#| eval: false

scatterplot <- function(df, x_var, y_var) {
  label <- rlang::englue("A scatterplot of _____ and _____, including a line of best fit.")
  
  df |> 
    ggplot(mapping = aes(x = _____, 
                         y = _____
                         )
           ) + 
    geom_point() + 
    geom_smooth(method = "lm", _____) +
    labs(title = _____)
}
```
