---
title: "Directories & Scientific Reproducibility"
editor: source
format: 
  revealjs:
    theme: ../style.scss
---

```{r}
#| include: false
#| label: packages

library(tidyverse)
```

# Reproducibility

# Reproducibility

> You should be able to send your project to someone else, and they can 
(re)produce the same plots / tables / results you did. 

## Principles of Reproducibility

- **Nuts and bolts**
  + [The documentation for a project should contain everything necessary (e.g., 
  scripts, Quarto files) to reproduce the results.]{.small}  


. . .

- **Portability**
  + [Any user should be able to run the scripts on their own computer or
  workspace.]{.small}

. . .

- **One-click Reproducibility**
  + [Reproducing all the computations should require (almost) nothing other than executing a single main script file.]{.small}


# How can a script I write run on someone else's computer???

# How can a script I write run on someone else's computer???

[Directories!]{.larger}

## What is a directory?

-   A **directory** is just a fancy name for a folder.

. . .

-   Your **working directory** is the folder that `R` "thinks" it lives in at
the moment.

```{r}
#| label: getwd
#| echo: true
#| eval: false
#| code-line-numbers: false

getwd()
```

```
"/Users/jcanner/Documents/STAT 210/videos/week-1"
```

## Paths

-   A **path** describes where a certain file or directory lives.

. . .

```{r}
#| label: getwd-2
#| echo: true
#| eval: false
#| code-line-numbers: false

getwd()
```

```
"/Users/jcanner/Documents/STAT 210/videos/week-1"
```

</br> 

. . .

This file lives on my computer in my user files (`/Users/`),

. . .

on my account (`/jcanner/`), 

. . .

in my Documents (`/Documents`), 

. . .

in a series of organized folders (`STAT 210/videos/week-1`). 

# Managing Your Class Directory

## File Management Check-in

::: {.incremental}
- Your `STAT210` folder should live in your **Documents**.

- It should have a `Week1` sub-folder. 

- Inside the `Week1` folder, there should be a `Lab1` sub-folder. 
:::
. . .

::: callout-caution
# Your folder cannot synch with anything online! 

Your STAT 210 folder **cannot** be in a folder stored on Google Drive, 0OneDrive, or iCloud!
Storing your project in this location will cause your code to periodically not
run and I cannot help you fix it.
:::
