---
title: "Lab 4: Importing NOAA Buoy Data from Monterey Bay"
author: "Instructions"
format: html
editor: source
number-sections: true
number-depth: 2
toc: true
code-tools: true
execute:
  echo: true
  eval: false
  message: false
  warning: false
  error: true
---

# Part 1: Setup

## GitHub Workflow

Set up your GitHub workflow (either using the method of creating a repository and using Version Control to set up your project or vice versa using the `usethis` package commands we have learned).

Use appropriate naming conventions for your project (see Code Style Guide), e.g. `lab-4-import-data`.    

Your project folder should contain the following:  

- .Rproj  
- [lab-4-student.qmd](../student/lab-4-student.qmd) 
- "data" folder
    + [buoy_2024.txt](../student/data/buoy_2024.txt)
- rendered document

You will submit a link to your GitHub repository with all content. 


## Seeking Help

Part of learning to program is learning from a variety of resources. Thus, I expect you will use resources that you find on the internet. There is, however, an important balance between copying someone else's code and *using their code to learn*. Therefore, if you use external resources, I want to know about it. 

- If you used Google, you are expected to "inform" me of any resources you used by **pasting the link to the resource in a code comment next to where you used that resource**.

- If you used ChatGPT, you are expected to "inform" me of the assistance you received by (1) indicating somewhere in the problem that you used ChatGPT (e.g., below the question prompt or as a code comment), and (2) downloading and attaching the `.txt` file containing your **entire** conversation with ChatGPT. ChatGPT can we used as a "search engine", but you should not copy and paste prompts from the lab or the code into your lab.  

Additionally, you are permitted and encouraged to work with your peers as you complete lab assignments, but **you are expected to do your own work**. Copying from each other is cheating, and letting people copy from you is also cheating. Please don't do either of those things.

## Lab Instructions

The questions in this lab are noted with numbers and boldface. Each question will require you to produce code, whether it is one line or multiple lines. 

This document is quite plain, meaning it does not have any special formatting. As part of your demonstration of creating professional looking Quarto documents, **I would encourage you to spice your documents up (e.g., declaring execution options, specifying how your figures should be output, formatting your code output, etc.).**

## Setup

In the code chunk below, load in the packages necessary for your analysis. You should only need the `tidyverse` package for this analysis.

```{r}
#| label: setup


```

# Part 2: Data Context
https://www.ndbc.noaa.gov/station_history.php?station=46092 

https://www.ndbc.noaa.gov/faq/measdes.shtml

Historical files have gone through post-processing analysis and represent the data sent to the archive centers. The formats for both are generally the same, with the major difference being the treatment of missing data. Missing data in the Historical files are variable number of 9's, depending on the data type (for example: 999.0 99.0).


## Reading the Data into `R`



```{r}
#| label: load-data

library(tidyverse)
buoy <- readr::read_table("../student/data/buoy_2024.txt", 
                          comment = "#",
                          col_types = cols(.default = col_character()),
                          na = c("99", "999"))
                          
write_csv(buoy2, "../student/data/buoy3.csv") 

buoy3 <- readr::read_csv("../student/data/buoy3.csv",
                          col_types = cols(.default = col_double())) |> 
  janitor::clean_names()


buoy3 |> 
  ggplot(mapping = aes(y = atmp, x = mm)) +
  geom_jitter(aes(color = mm), alpha = 0.1, width = 0.1) +
  geom_smooth(color = "black") + 
  guides(fill = "none", color = "none") +
  scale_fill_viridis_c() + 
  scale_color_viridis_c() +
  theme_light() +
  scale_x_continuous(breaks = c(1:12)) +
  labs(x = "Month in 2024",
       y = "Air Temperature (Celsius)",
       title = "Monterey Bay Air Temperature in 2024",
       subtitle = "Station 46092 Buoy Measurements")
  

```



# Part ##: Find a Data Set
For your final portfolio, you will choose a data set to clean and provide an analysis of through visualization (or more if you have the experience).  For this week, choose a data set

- It is a real-world application of R that has not exactly been worked out before (e.g. it isn't a demo from some package or blog).  
- It is interesting to you.  
- It involves data and analyzing or presenting that data. The data may be data you have from a lab, or something you have retrieved from the web, some examples of good sources:  FBI database on crime statistics, National Oceanic and Atmospheric Administration, World Health Organization, Twitter, Yahoo finance data, etc.  If you are having problems finding a dataset, see the resources at the end of the project description.  
- The analysis and presentation is useful in the real-world.  

One fun source for data is [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday) which posts a new data set almost every week. It is a great place to find ideas and sources. 

**For your found data write a brief description of the data and and provide its source.** 



# Lab 4 Submission

For Lab 4 you will submit the link to your GitHub repository. Your rendered file is required to have the following specifications in the YAML options (at the top of your document):

-   have the plots embedded (`embed-resources: true`)
-   include your source code (`code-tools: true`)
-   include all your code and output (`echo: true`)

**If any of the options are not included, your Lab 3 or Challenge 3 assignment will receive an "Incomplete" and you will be required to submit a revision.**

In addition, your document should not have any warnings or messages output in your HTML document. **If your HTML contains warnings or messages, you will receive an "Incomplete" for document formatting and you will be required to submit a revision.**
